Reproducible Research - Assignment 1
====================================

## Loading and Processing the data

Unzipping and reading the .csv file. First set the present working directory to where the downloaded file is located, using setwd() function.

```{r file, echo=TRUE}
activity <- unzip('repdata_data_activity.zip')
data <- read.csv("activity.csv", header = TRUE)
```

## Mean total number of steps taken per day

Total number of steps taken per day

```{r steps/day, include = TRUE}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, sum, na.rm = TRUE)
```

Histogram of the total number of steps taken each day

```{r histogram of total steps per day, include = TRUE}
qplot(total.steps, binwidth = 1000, xlab = "Total number of steps taken each day", ylab = "Count")
```

Mean and median of the total number of steps taken per day

```{r mean and median of total steps per day, include = TRUE}
mean(total.steps, na.rm = TRUE)
median(total.steps, na.rm = TRUE)
```

## Average daily activity pattern

Time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r Average daily activity pattern, include = TRUE}
library(ggplot2)
average <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval), mean, na.rm = TRUE)
ggplot(average, aes(interval, steps)) + geom_line() + xlab("5-minutes interval") + ylab("average number of steps taken")
```

The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps is

```{r , include = TRUE}
average[which.max(average$steps), ]
```

## Imputing missing data

There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r , include = TRUE}
missing_values <- is.na(data$steps)
table(missing_values)
```

All the missing values will be filled with the mean value of 5- minutes interval

```{r filling missing values, include=TRUE}
stepsaverage <- aggregate(steps ~ interval, data, mean)
filling <- numeric()
for (i in 1:nrow(data)) {
  obs <- data[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(stepsaverage, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  filling <- c(filling, steps)
}
```

New data that is equal to the original dataset with missing data filled in.

```{r new data set}
new_data <- data
new_data$steps <- filling
```

Histogram of the total number of steps taken each day

```{r histogram, include = TRUE}
stepsperday <- tapply(new_data$steps, new_data$date, sum)
qplot(stepsperday, binwidth = 500, xlab = "Total Steps per day", ylab = "Frequency using binwidth 500")
```

Calculate and report the mean and median total number of steps taken per day.

```{r mean and median, include = TRUE}
mean(stepsperday)
median(stepsperday)
```

Does the mean and median values of imputed data differ from the estimates from the first part of the assignment? Also, what is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r Mean and Median Impact - Imputed vs Non-Imputed, include = TRUE}
mean(stepsperday) - mean(total.steps, na.rm = TRUE)
median(stepsperday) - median(total.steps, na.rm = TRUE)
```

Yes, the mean and median differ(is higher) after imputing. The impact is calculated as above. 

## Difference in activity patterns between weekdays and weekends

Creating a new factor variable in the dataset with two levels ??? ???weekday??? and ???weekend??? indicating whether a given date is a weekday or weekend day.

```{r Activity pattern diff weekends vs weekdays, include = TRUE}
new_data$date <- as.Date(as.factor(new_data$date))
day <- weekdays(new_data$date)
daytype <- vector()
for(i in 1:nrow(new_data)) {
  if(day[i] == "Saturday") {
    daytype[i] <- "Weekend"
  } else if(day[i] == "Sunday"){
    daytype[i] <- "Weekend"
  } else {
    daytype[i] <- "Weekday"
  }
}
```

Panel plot containing a time series plot (i.e. type = "l"\color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r Panel plot, include = TRUE}
stepsbyday <- aggregate(steps ~ interval + daytype, new_data, mean)
ggplot(stepsbyday, aes(interval,steps)) + geom_line() + facet_grid(daytype ~ .) + xlab("5-minute interval") + ylab("Number of Steps")
```
